---
title: "Creating a Transect Line"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, warning=FALSE, results="hide", message=FALSE, cache=TRUE, fig.align="center")
```

The goal is to set up a transect line with evenly spaced points defined by latitude and longitude. We will use this transect to download ocean remote-sensing variables at those locations.

First we will set up our two points.

```{r}
library(data.table)
df <- data.frame(
  Latitude=c(9.9413972, 9.2),
  Longitude=c(76.2, 74.5)
  )
data.table::setDT(df)
```

Here is where they are on a map.

```{r}
library(rworldmap)
newmap <- rworldmap::getMap(resolution = "low")
plot(newmap, xlim = c(75,76), ylim = c(8,13), asp = 1)
points(df$Longitude, df$Latitude, col = "red", cex = .6)
```

Next we will interpolate points every 10km on this line.
```{r}
dist.between = 10000 #in meters

library(geotools)
library(geosphere)    # for distHaversine
get.dist <- function(lon, lat) geosphere::distHaversine(tail(cbind(lon,lat),-1), head(cbind(lon,lat),-1))
# The := function is in data.table
df[,dist:=c(0,cumsum(get.dist(Longitude,Latitude)))]

interp.dist <- function(var,dist) approx(dist,var,xout=seq(min(dist),max(dist),by=dist.between))$y
result <- setDT(df)[,lapply(.SD,interp.dist,dist=dist), 
                    .SDcols=c("Latitude","Longitude","dist")]
```

Now we can plot these points.

```{r}
# plot the result
newmap <- rworldmap::getMap(resolution = "low")
plot(newmap, xlim = c(74,76), ylim = c(8,13))
points(df$Longitude, df$Latitude, col = "red", cex = .6)
points(Latitude~Longitude,result, col="blue")
axis(side=1)
axis(side=2)
```

We can plot with `ggplot()` also.

```{r}
library(ggplot2)
india <- rworldmap::getMap(resolution = "low")

(with_india <- ggplot() +
	geom_path(data = india, 
		aes(x = long, y = lat, group = group),
		fill = NA, colour = "black") + 
	geom_point(data = result, 
		aes(x = Longitude, y = Latitude)) +
  xlim(74,80) +
  ylim(7,13) +
	coord_quickmap() +  # Prevents stretching when resizing
	theme_classic() +  # Remove ugly grey background
	xlab("Longitude") +
	ylab("Latitude"))
```

That's really ugly so let's improve it.

Download bathymetry data from NOAA.

```{r}
library(marmap)#if it is not in your package list install it simply by marmap on the dialog box
xlims <- c(73,80)
ylims <- c(6,15)
bathydata<- getNOAA.bathy(xlims[1], xlims[2], ylims[1], ylims[2])#Be sure the locations are same
```

Plot the data with our transect line. You have to tweak the plot region otherwise you will have white space around your plot. It has to do with setting `asp=1` (default) to not distort your plot. You can change to `asp=NA` once your plot size is close to `asp=1`.

```{r fig.width=5.2, fig.height=7}
# code adapted from 
# https://pepprbook.com/2018/02/17/mapping-in-r/
#Load the bathymetry data from NOAA using marmap
library(rworldmap)
library(rworldxtra)
library(maps)
#Make some colour combinations.Here I took blues and greys

blues <- c("lightsteelblue4", "lightsteelblue3",
 "lightsteelblue2", "lightsteelblue1")
greys <- c(grey(0.6), grey(0.93), grey(0.99))
#Now plot the data 'bathydata' with the colours we created
#par(mar=c(5, 10, 5, 10) + 0.1)
plot(bathydata, image = TRUE, land = TRUE, lwd = 0.03,
 bpal = list(c(0, max(bathydata), greys),
 c(min(bathydata), 0, blues)), asp=1)
#make thickened contour lines
plot(bathydata, n = 4, lwd = 0.4, add = TRUE)
#Add the map overlay
map("world","India",col = "grey",fill = TRUE,add=TRUE,xlim=xlims,ylim=ylims, lwd=2)
#Add your sampling points(pch=25 makes it inverted triangles)
points(result$Longitude,result$Latitude,pch=16,col="black",cex=.5)
box(lty="solid", lwd=2)
```

We can also make the plot with `ggplot()`.

```{r}
autoplot(bathydata, geom=c("r", "c")) + 
  scale_fill_etopo() +
  ylim(7,13) +
  xlim(73,79)+
  geom_point(aes(x=Longitude, y=Latitude), data=result, alpha=1)
```